Datasheet for 'NuSEDS-New Salmon Escapement Database System

### 2024-04-19 other/datasheets/datasheet_nuseds.qmd
complete datasheets

---
title: "Datasheet for 'Pacific Region Commercial Salmon Fishery In-season Catch Estimates'"
author: 
  - Timothius Prajogi
date: today
date-format: long
format: pdf
number-sections: true
bibliography: references.bib
---

@@

1. *For what purpose was the dataset created? Was there a specific task in mind? Was there a specific gap that needed to be filled? Please provide a description.*
    - The database was built in 2001 and continues to be the current repository for all data pertaining to DFO’s Pacific Region Commercial Salmon Logbook program.
2. *Who created the dataset (for example, which team, research group) and on behalf of which entity (for example, company, institution, organization)?*
    - Fisheries and Oceans Canada
3. *Who funded the creation of the dataset? If there is an associated grant, please provide the name of the grantor and the grant name and number.*
    - Government funded.
4. *Any other comments?*
    - No.

@@

	- The instances of the dataset represent catch estimates for all types of salmon, the amount of vessels that were licensed, the amount of days that were fished by boats in total, and the year and area this took place in. They are separated by the type of fishing that was done.

@@

	- 238 observations for gill nets, 190 for seine nets, 468 for trolling vessels.

@@

	- It contains all recoreded observations. 

@@

	- It consists of total count values of each feature.

@@

	- The rows are labeled by the fishery and licence area.

@@

	- Some observations have been redacted due to privacy reasons.

@@

@@

	- Self contained.
	- Confidential data has been removed.

@@

	- It identifies salmon by their genus/species.

@@

1. *How was the data associated with each instance acquired? Was the data directly observable (for example, raw text, movie ratings), reported by subjects (for example, survey responses), or indirectly inferred/derived from other data (for example, part-of-speech tags, model-based guesses for age or language)? If the data was reported by subjects or indirectly inferred/derived from other data, was the data validated/verified? If so, please describe how.*
	- Fishers were instructed to log their catches, and total estimates were made by Fisheries and Oceans Canada based on those logs.
2. *What mechanisms or procedures were used to collect the data (for example, hardware apparatuses or sensors, manual human curation, software programs, software APIs)? How were these mechanisms or procedures validated?*
	- Human curation with the assistance of programs helped create these estimates.
3. *If the dataset is a sample from a larger set, what was the sampling strategy (for example, deterministic, probabilistic with specific sampling probabilities)?*
	- No.
4. *Who was involved in the data collection process (for example, students, crowdworkers, contractors) and how were they compensated (for example, how much were crowdworkers paid)?*
	- Both the fishers and the researchers were involved in creating this dataset.
5. *Over what timeframe was the data collected? Does this timeframe match the creation timeframe of the data associated with the instances (for example, recent crawl of old news articles)? If not, please describe the timeframe in which the data associated with the instances was created.*
	- From 2005 - 2022.
6. *Were any ethical review processes conducted (for example, by an institutional review board)? If so, please provide a description of these review processes, including the outcomes, as well as a link or other access point to any supporting documentation.*
	- Not stated.
7. *Did you collect the data from the individuals in question directly, or obtain it via third parties or other sources (for example, websites)?*
	- I collected directly from the source.
8. *Were the individuals in question notified about the data collection? If so, please describe (or show with screenshots or other information) how notice was provided, and provide a link or other access point to, or otherwise reproduce, the exact language of the notification itself.*
	- No.
9. *Did the individuals in question consent to the collection and use of their data? If so, please describe (or show with screenshots or other information) how consent was requested and provided, and provide a link or other access point to, or otherwise reproduce, the exact language to which the individuals consented.*
	- No they were not, they had licensed the dataset to not require such authorization.
10. *If consent was obtained, were the consenting individuals provided with a mechanism to revoke their consent in the future or for certain uses? If so, please provide a description, as well as a link or other access point to the mechanism (if appropriate).*
	- N/A
11. *Has an analysis of the potential impact of the dataset and its use on data subjects (for example, a data protection impact analysis) been conducted? If so, please provide a description of this analysis, including the outcomes, as well as a link or other access point to any supporting documentation.*
	- Not stated.
12. *Any other comments?*
	- No.

@@

1. *Was any preprocessing/cleaning/labeling of the data done (for example, discretization or bucketing, tokenization, part-of-speech tagging, SIFT feature extraction, removal of instances, processing of missing values)? If so, please provide a description. If not, you may skip the remaining questions in this section.*
	- The rows ommitted for privacy reasons were removed and skipped.
2. *Was the "raw" data saved in addition to the preprocessed/cleaned/labeled data (for example, to support unanticipated future uses)? If so, please provide a link or other access point to the "raw" data.*
	- Yes, the raw data in this project can be found in `input/data` and can be found from the source here: https://open.canada.ca/data/en/dataset/c48669a3-045b-400d-b730-48aafe8c5ee6.
3. *Is the software that was used to preprocess/clean/label the data available? If so, please provide a link or other access point.*
	- Yes, R a statistical programming language was used and the documentation can be found here: https://www.r-project.org/
4. *Any other comments?*
	- No.

@@

1. *Has the dataset been used for any tasks already? If so, please provide a description.*
	- Not stated.
2. *Is there a repository that links to any or all papers or systems that use the dataset? If so, please provide a link or other access point.*
	- No.
3. *What (other) tasks could the dataset be used for?*
	- The dataset could be used to monitor the fishing industry, its success and failures as well as its impact on other ecosystems and environments.
4. *Is there anything about the composition of the dataset or the way it was collected and preprocessed/cleaned/labeled that might impact future uses? For example, is there anything that a dataset consumer might need to know to avoid uses that could result in unfair treatment of individuals or groups (for example, stereotyping, quality of service issues) or other risks or harms (for example, legal risks, financial harms)? If so, please provide a description. Is there anything a dataset consumer could do to mitigate these risks or harms?*
	- No.
5. *Are there tasks for which the dataset should not be used? If so, please provide a description.*
	- It is to be noted that this dataset is only for the Canadian Pacific, and should only be used in analyses that are about that region.
6. *Any other comments?*
	- No.

@@

1. *Who will be supporting/hosting/maintaining the dataset?*
	- Maintenance is done by the publisher Fisheries and Oceans Canada.
2. *How can the owner/curator/manager of the dataset be contacted (for example, email address)?*
	- They have provided an email address to be contacted: shelee.hamilton@dfo-mpo.gc.ca
3. *Is there an erratum? If so, please provide a link or other access point.*
	- No.
4. *Will the dataset be updated (for example, to correct labeling errors, add new instances, delete instances)? If so, please describe how often, by whom, and how updates will be communicated to dataset consumers (for example, mailing list, GitHub)?*
	- Not stated.
5. *If the dataset relates to people, are there applicable limits on the retention of the data associated with the instances (for example, were the individuals in question told that their data would be retained for a fixed period of time and then deleted)? If so, please describe these limits and explain how they will be enforced.*
	- N/A
6. *Will older versions of the dataset continue to be supported/hosted/maintained? If so, please describe how. If not, please describe how its obsolescence will be communicated to dataset consumers.*
	- Not stated.
7. *If others want to extend/augment/build on/contribute to the dataset, is there a mechanism for them to do so? If so, please provide a description. Will these contributions be validated/verified? If so, please describe how. If not, why not? Is there a process for communicating/distributing these contributions to dataset consumers? If so, please provide a description.*
	- Not stated.
8. *Any other comments?*
	- No.

>>>>

---
title: "Datasheet for 'NuSEDS-New Salmon Escapement Database System'"
author: 
  - Timothius Prajogi
date: today
date-format: long
format: pdf
number-sections: true
bibliography: references.bib
---

@@

1. *For what purpose was the dataset created? Was there a specific task in mind? Was there a specific gap that needed to be filled? Please provide a description.*
    - The dataset was created to keep track and monitor salmon spawning records and spawner abundance estimates. It is the primary database used to track salmon for Fisheries and Oceans Canada and it enables future research. It reports spawning observations for 9100_ individual populations. 
2. *Who created the dataset (for example, which team, research group) and on behalf of which entity (for example, company, institution, organization)?*
    - The dataset was published by Fisheries and Oceans Canada.
3. *Who funded the creation of the dataset? If there is an associated grant, please provide the name of the grantor and the grant name and number.*
    - This is a government led program.
4. *Any other comments?*
    - No.

@@

	- An instance in a dataset represents the spawning salmon population observed in a specific waterbody during a specific year.

@@

	- There are 413,434 observations made.

@@

	- The dataset contains all recorded observations made in the Pacific waterbodies of Canada, and is intended to be a comprehensive dataset.

@@

	- Each observation notes down amount of salmon observed, when it was observed, the type of salmon, any broodstock removed, and the waterbody the observation was made in. 

@@

	- Each observation is labeled with a unique identifier labeled `ACT_ID`.

@@

	- The instance may or may not record certain data regarding the spawning population. For broodstock data, when it is not present it signifies no broodstock was removed from the waterway. Specific date values are sometimes missing, being insignificant data as spawning season generally comes around at the same time, with specific dates recorded only for certain waterbodies.

@@

	- The waterbodies and regions are combined to create a unique id which can be used to reference all other observations made about that waterbody and region for any year, species, or otherwise.
	- No datasplits were recommended.
	- Historical data that was imported was included but is largely blank. This data just recorded basic observations, with newer observations provided much more detail.
	- The dataset is self-contained, hosted on Canada's open data portal.

@@

@@

	- As mentioned the observations were split by species, waterbody, and region. These attributes are identified in their respective columns.

@@

1. *How was the data associated with each instance acquired? Was the data directly observable (for example, raw text, movie ratings), reported by subjects (for example, survey responses), or indirectly inferred/derived from other data (for example, part-of-speech tags, model-based guesses for age or language)? If the data was reported by subjects or indirectly inferred/derived from other data, was the data validated/verified? If so, please describe how.*
	- The data was collected as a survey, made from observations gathered from field researchers. Estimates were made as to the exact counts, with the scale of the matter being so large and spanning long periods.
2. *What mechanisms or procedures were used to collect the data (for example, hardware apparatuses or sensors, manual human curation, software programs, software APIs)? How were these mechanisms or procedures validated?*
	- Human curation was necessary to gather the data, along with measurement instruments.
3. *If the dataset is a sample from a larger set, what was the sampling strategy (for example, deterministic, probabilistic with specific sampling probabilities)?*
	- N/A
4. *Who was involved in the data collection process (for example, students, crowdworkers, contractors) and how were they compensated (for example, how much were crowdworkers paid)?*
	- This is not explicitly mentioned, however the data was provided by Fisheries and Oceans Canada.
5. *Over what timeframe was the data collected? Does this timeframe match the creation timeframe of the data associated with the instances (for example, recent crawl of old news articles)? If not, please describe the timeframe in which the data associated with the instances was created.*
	- The data is collected yearly, and has been collected since 1950 with the last update being 2022. The dataset is updated quarterly as finalized observations come in.
6. *Were any ethical review processes conducted (for example, by an institutional review board)? If so, please provide a description of these review processes, including the outcomes, as well as a link or other access point to any supporting documentation.*
	- No ethical review processes were stated.
7. *Did you collect the data from the individuals in question directly, or obtain it via third parties or other sources (for example, websites)?*
	- I collected the data directly from their API source.
8. *Were the individuals in question notified about the data collection? If so, please describe (or show with screenshots or other information) how notice was provided, and provide a link or other access point to, or otherwise reproduce, the exact language of the notification itself.*
	- No they were not, they had licensed the dataset to not require such authorization.
9. *Did the individuals in question consent to the collection and use of their data? If so, please describe (or show with screenshots or other information) how consent was requested and provided, and provide a link or other access point to, or otherwise reproduce, the exact language to which the individuals consented.*
	- N/A
10. *If consent was obtained, were the consenting individuals provided with a mechanism to revoke their consent in the future or for certain uses? If so, please provide a description, as well as a link or other access point to the mechanism (if appropriate).*
	- N/A
11. *Has an analysis of the potential impact of the dataset and its use on data subjects (for example, a data protection impact analysis) been conducted? If so, please provide a description of this analysis, including the outcomes, as well as a link or other access point to any supporting documentation.*
	- No analysis was stated.
12. *Any other comments?*
	- The background review process for this dataset was not stated, being a government database.

@@

1. *Was any preprocessing/cleaning/labeling of the data done (for example, discretization or bucketing, tokenization, part-of-speech tagging, SIFT feature extraction, removal of instances, processing of missing values)? If so, please provide a description. If not, you may skip the remaining questions in this section.*
	- Yes, N/A values needed to be handled, either treated as 0, or their observations ignored. Depending on the requirements, salmon species were grouped together by year to get an understanding of the grander scale of the population.
2. *Was the "raw" data saved in addition to the preprocessed/cleaned/labeled data (for example, to support unanticipated future uses)? If so, please provide a link or other access point to the "raw" data.*
	- Yes, the raw data in this project can be found in `input/data` and can be found from the source here: https://open.canada.ca/data/en/dataset/c48669a3-045b-400d-b730-48aafe8c5ee6.
3. *Is the software that was used to preprocess/clean/label the data available? If so, please provide a link or other access point.*
	- Yes, R a statistical programming language was used and the documentation can be found here: https://www.r-project.org/
4. *Any other comments?*
	- No.

@@

1. *Has the dataset been used for any tasks already? If so, please provide a description.*
	- I imagine so, but usage has not been documented.
2. *Is there a repository that links to any or all papers or systems that use the dataset? If so, please provide a link or other access point.*
	- No.
3. *What (other) tasks could the dataset be used for?*
	- The dataset covers a lot of features that can be used in various aspects of salmon health monitoring. Other research surrounding each of the different features, spawn timings, broodstock, population, can be done.
4. *Is there anything about the composition of the dataset or the way it was collected and preprocessed/cleaned/labeled that might impact future uses? For example, is there anything that a dataset consumer might need to know to avoid uses that could result in unfair treatment of individuals or groups (for example, stereotyping, quality of service issues) or other risks or harms (for example, legal risks, financial harms)? If so, please provide a description. Is there anything a dataset consumer could do to mitigate these risks or harms?*
	- No.
5. *Are there tasks for which the dataset should not be used? If so, please provide a description.*
	- It is to be noted that this dataset is only for the Canadian Pacific, and should only be used in analyses that are about that region.
6. *Any other comments?*
	- No.

@@

1. *Who will be supporting/hosting/maintaining the dataset?*
	- Maintenance is done by the publisher Fisheries and Oceans Canada.
2. *How can the owner/curator/manager of the dataset be contacted (for example, email address)?*
	- They have provided an email address to be contacted: Zhipeng.Wu@dfo-mpo.gc.ca
3. *Is there an erratum? If so, please provide a link or other access point.*
	- No.
4. *Will the dataset be updated (for example, to correct labeling errors, add new instances, delete instances)? If so, please describe how often, by whom, and how updates will be communicated to dataset consumers (for example, mailing list, GitHub)?*
	- Not stated.
5. *If the dataset relates to people, are there applicable limits on the retention of the data associated with the instances (for example, were the individuals in question told that their data would be retained for a fixed period of time and then deleted)? If so, please describe these limits and explain how they will be enforced.*
	- N/A
6. *Will older versions of the dataset continue to be supported/hosted/maintained? If so, please describe how. If not, please describe how its obsolescence will be communicated to dataset consumers.*
	- Not stated.
7. *If others want to extend/augment/build on/contribute to the dataset, is there a mechanism for them to do so? If so, please provide a description. Will these contributions be validated/verified? If so, please describe how. If not, why not? Is there a process for communicating/distributing these contributions to dataset consumers? If so, please provide a description.*
	- Not stated.
8. *Any other comments?*
	- No.
